---
title: "All_Species_Dashboard"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    horizontal_layout: fill
    theme: default
---

```{r setup, include=FALSE}
require(tidyverse)
require(data.table)
require(shiny)
require(here)
require(sp)
require(rgdal)
require(leaflet)
require(leafem)
require(leaflet.extras)
require(flexdashboard)
require(inborutils)
require(imager)
require(RCurl)
require(rsconnect)

if("trias" %in% rownames(installed.packages()) == FALSE) {
  devtools::install_github("trias-project/trias")}

require(trias)

minjaar <- lubridate::year(Sys.Date())-5
maxjaar <- lubridate::year(Sys.Date())-2

bioregions <- c("Atlantic", "Continental")
ts <- c("t0", "t1")
batches <- c("Batch1", "Batch2", "Batch3")

crs_wgs <- CRS("+proj=longlat +datum=WGS84 +no_defs ")
crs_bel <- CRS("+proj=lcc +lat_1=51.16666723333333 +lat_2=49.8333339 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.013 +y_0=5400088.438 +ellps=intl +towgs84=-106.869,52.2978,-103.724,0.3366,-0.457,1.8422,-1.2747 +units=m +no_defs ")
```

```{r data, include=FALSE, context="data", cache=TRUE}
bioreg_bel_clip <- readOGR("https://github.com/inbo/IAS_Species_Dashboard/raw/3_version1.0/Data/Spatial/bioreg_bel_clip.geojson", "bioreg_bel_clip", stringsAsFactors = FALSE)
bel_borders <- readOGR("https://github.com/inbo/IAS_Species_Dashboard/raw/3_version1.0/Data/Spatial/Belgie.geojson", "Belgie", stringsAsFactors = FALSE)

proj4string(bel_borders) <- crs_bel

grid_pal <- colorFactor(palette = c("red", "black"), domain = utm_polygons$grid, levels = c("utm1", "utm5"))
bioreg_pal <- colorFactor(palette = c("darkgrey", "white"), domain = bioreg_bel_clip$BIOGEO, levels = c("Continental", "Atlantic"))

basemap <- leaflet(bioreg_bel_clip) %>% 
  addTiles() %>% 
  addPolygons(fillColor = ~bioreg_pal(BIOGEO),
              fillOpacity = 0.5,
              stroke = FALSE) %>% 
  addPolylines(data = spTransform(bel_borders, crs_wgs),
               color = "black",
               opacity = 1,
               weight = 2) %>% 
  addScaleBar(position = "bottomleft") %>% 
  addLogo(img = "https://cdn2.iconfinder.com/data/icons/map-and-navigation-12/48/57-512.png", src = "remote", position = "bottomright") %>% 
  setMapWidgetStyle(list(background= "white"))


utm_merged <-  readOGR("https://github.com/inbo/IAS_Species_Dashboard/raw/3_version1.0/Data/Spatial/utm_merged.geojson",
                       "utm_merged", stringsAsFactors = FALSE)

stats_merged <- read_delim("https://github.com/inbo/IAS_Species_Dashboard/raw/3_version1.0/Data/Interim/stats_merged.csv", ";")

spec_ind <- read_delim("https://raw.githubusercontent.com/inbo/IAS_Species_Dashboard/master/Data/Lists/nameserver_species_identifiers_for_t0.csv", 
                       ";", escape_double = FALSE, col_types = cols(t0_from = col_date(format = "%d/%m/%Y"), 
                                                                    t0_to = col_date(format = "%d/%m/%Y"), 
                                                                    t1_from = col_date(format = "%d/%m/%Y"), 
                                                                    t1_to = col_date(format = "%d/%m/%Y")), 
                       trim_ws = TRUE, locale = locale("en"))

species_choices <- spec_ind %>% 
  filter(!is.na(Batch)) %>% 
  mutate(Species = case_when(grepl("Salvinia", gbifapi_acceptedScientificName) ~ "Salvinia auriculata Aubl.",
                             TRUE ~ gbifapi_acceptedScientificName)) %>% 
  distinct(Species) 

species_choices$Species <- str_sort(species_choices$Species)

spec_ind <- spec_ind %>% 
  filter(!is.na(Batch))

df_ts_compact <- read_tsv("https://github.com/inbo/IAS_Species_Dashboard/raw/3_version1.0/Data/Interim/df_timeseries_redux.tsv")
```

Sidebar {.sidebar}
==================================
### Species dropdown
Select a species from the dropdown
All graphs, maps and tables will be rendered
for the selected species.

```{r Species dropdown, context="server"}
selectInput(inputId = "species", label = NULL, choices = species_choices, selected = species_choices[1], multiple = FALSE, selectize = FALSE)
```

```{r Inclusion info, context="server"}
renderText({print("")})
renderText({
  spec_ind_sub <- spec_ind %>% 
    filter(gbifapi_acceptedScientificName == input$species) %>% 
    distinct(gbifapi_acceptedKey, Batch, t0_from, t0_to)
  
  if(spec_ind_sub$Batch == 1){
    batch <- paste0(spec_ind_sub$Batch, "st")
  }
  if(spec_ind_sub$Batch == 2){
    batch <- paste0(spec_ind_sub$Batch, "nd")
  }
  if(spec_ind_sub$Batch == 3){
    batch <- paste0(spec_ind_sub$Batch, "rd")
  }
  if(spec_ind_sub$Batch > 3 ){
    batch <- paste0(spec_ind_sub$Batch, "th")
  }
  
  text_general <- paste0(input$species, " was included into the European list with the ", batch, " batch on ",  spec_ind_sub$t0_to, ".") 
  print(text_general)
})
renderText({print("")})
```


### Emergence evaluation period
The emergence status for the species will
be assessed for this period

```{r eval_years slider, context="server"}
sliderInput(inputId = "evaluation_years", label = NULL, min = minjaar, max = lubridate::year(Sys.Date()), value = c(maxjaar-2,maxjaar), step = 1, sep = "", dragRange = TRUE)
```

```{r em_status info, context="server"}
renderText({
  if(max(input$evaluation_years) > maxjaar){
    print("WARNING: Due to reporting lag, it is possible the emergence status is not calculated correctly!")
  }else{
    print("")
  }
})

renderText({print("")})

renderText({
  spec_ind_sub <- spec_ind %>% 
    filter(gbifapi_acceptedScientificName == input$species)
  
  df_key <- df_ts_compact %>% 
    filter(taxonKey %in% spec_ind_sub$gbifapi_acceptedKey)
  
  em_decision_rules_occs_BE <- data.frame(taxonKey = 0, year = 0, em_status = 1, dr_1 = FALSE, dr_2 = FALSE,dr_3 = FALSE,dr_4 = FALSE)
  
  if(nrow(df_key) > 0){
    for(j in input$evaluation_years){
      temp <- apply_decision_rules(
        df = df_key,
        y_var = "obs",
        eval_year = j,
        year = "year"
      )
      em_decision_rules_occs_BE <- rbind(em_decision_rules_occs_BE, temp)
    }
  }
  
  em_status_code <- max(em_decision_rules_occs_BE$em_status, na.rm = TRUE)
  
  if(em_status_code == 0){
    em_status <- "not emerging"
  }
  if(em_status_code == 1){
    em_status <- "unclear"
  }
  if(em_status_code == 2){
    em_status <- "potentially emerging"
  }
  if(em_status_code == 3){
    em_status <- "emerging"
  }
  
  em_status_text <- paste0("The emergence status of ", input$species, " was considered ",em_status, " in Belgium between ",  min(input$evaluation_years), " & ", max(input$evaluation_years), ".") 
  
  alt_em_status_text <- paste0("The emergence status of ", input$species, " between ",  min(input$evaluation_years), " & ", max(input$evaluation_years), " cannot be assessed.") 
  
  if(nrow(df_key) > 0){
    print(em_status_text)
  }else{
    print(alt_em_status_text)
  }
  
})
```

T0 - Baseline
==========================================

Row {data-height=50}
----------------------------------

```{r echo=FALSE, context="server"}
renderText({
  spec_ind_sub <- spec_ind %>% 
    filter(gbifapi_acceptedScientificName == input$species) %>% 
    distinct(gbifapi_acceptedKey, Batch, t0_from, t0_to)
  
  text_t0 <- paste0("The graphs, tables and maps below contain observations from ", spec_ind_sub$t0_from, " to ",  spec_ind_sub$t0_to,".") 
  print(text_t0)
  
})
```

Row {data-height=400}
----------------------------------

```{r, context="server"}
renderLeaflet({
  # force(basemap)
  # force(utm_merged)
  
  utm_t0 <- subset(utm_merged, utm_merged$tx == "t0")
  spec_ind_sub <- spec_ind %>% 
    filter(gbifapi_acceptedScientificName == input$species) %>% 
    distinct(gbifapi_acceptedKey, Batch, t0_from, t0_to)
  if(!is.na(spec_ind_sub$t0_from)){
    utm_spec <- subset(utm_t0, utm_t0$species == input$species)
    
    t0_map <- basemap %>% 
      addPolygons(data = utm_spec,
                  color = ~grid_pal(grid),
                  fill = FALSE,
                  opacity = 1,
                  weight = 1) %>% 
      addLegend(data = utm_spec,
                pal = grid_pal, 
                values = ~grid,
                position = "topright",
                title = "legenda")
    
    print(t0_map)
  }
})
```

Row {data-height=300}
----------------------------------

### Stats

*SAC: Special Areas of Conservation comprise of areas determined under the Habitat Directive

```{r, context="server"}
renderTable({
  spec_stats <- stats_merged %>% 
    filter(gbfp_SN == input$species & tx == "t0") %>% 
    select(-tx)
  
  bioregion_test <- spec_stats %>% 
    filter(bioregion %in% bioregions) 
  
  if(nrow(bioregion_test) < 2){
    if(nrow(bioregion_test) > 0){
      missing_bioreg <- subset(bioregions, !grepl(pattern = bioregion_test$bioregion, bioregions))
      spec_stats <- spec_stats %>% 
        ungroup() %>% 
        add_row(gbfp_SN = input$species, bioregion = missing_bioreg, utm1 = NA, utm2 = NA, utm5 = NA, utm10 = NA, sac_perc = NA)
    }else{
      for(bio in bioregions){
        spec_stats <- spec_stats %>% 
          ungroup() %>% 
          add_row(gbfp_SN = input$species, bioregion = bio, utm1 = NA, utm2 = NA, utm5 = NA, utm10 = NA, sac_perc = NA)
      }
    }
  }
  
  temp_stats_3 <- spec_stats  %>% 
    ungroup() %>% 
    filter(!is.na(bioregion)) %>% 
    select(-gbfp_SN) %>% 
    rename(`UTM 1x1km` = utm1,
           `UTM 2x2km` = utm2,
           `UTM 5x5km` = utm5,
           `UTM 10x10km` = utm10,
           `% UTM 1x1km squares with SACs*` = sac_perc) %>% 
    t()
  
  colnames(temp_stats_3) <- temp_stats_3[1,]
  temp_stats_3 <- temp_stats_3[-1,]
  temp_stats_3 <- as.data.frame(temp_stats_3) %>% 
    rownames_to_column(var = "grid") 
  print(temp_stats_3)
})
```



T1 - Report
==================================

Row {data-height=50}
----------------------------------

```{r eval=TRUE, context="server"}
renderText({
  spec_ind_sub <- spec_ind %>% 
    filter(gbifapi_acceptedScientificName == input$species) %>% 
    distinct(gbifapi_acceptedKey, Batch, t1_from, t1_to)
  if(!is.na(spec_ind_sub$t1_from)){
    text_t1 <- paste0("The graphs, tables and maps below contain observations from ", spec_ind_sub$t1_from, " to ",  spec_ind_sub$t1_to,".") 
    print(text_t1)
  }else{
    text_t1_alt <- paste0(input$species, " hasn't been the subject of a t1 reporting cycle")
    print(text_t1_alt)
  }
})
```

Row {data-height=650}
----------------------------------

```{r, context="server"}
renderLeaflet({
  
  # force(basemap)
  # force(utm_merged)
  
  utm_t1 <- subset(utm_merged, utm_merged$tx == "t1")
  spec_ind_sub <- spec_ind %>% 
    filter(gbifapi_acceptedScientificName == input$species) %>% 
    distinct(gbifapi_acceptedKey, Batch, t1_from, t1_to)
  if(!is.na(spec_ind_sub$t1_from)){
    utm_spec <- subset(utm_t1, utm_t1$species == input$species)
    
    t1_map <- basemap %>% 
      addPolygons(data = utm_spec,
                  color = ~grid_pal(grid),
                  fill = FALSE,
                  opacity = 1,
                  weight = 1) %>% 
      addLegend(data = utm_spec,
                pal = grid_pal, 
                values = ~grid,
                position = "topright",
                title = "legenda")
    
    print(t1_map)
  }
})
```

Row {data-height=350}
----------------------------------

### Stats 

*SAC: Special Areas of Conservation comprise of areas determined under the Habitat Directive

```{r, context="server"}
renderTable({
  spec_stats <- stats_merged %>% 
    filter(gbfp_SN == input$species & tx == "t1") %>% 
    select(-tx)
  
  bioregion_test <- spec_stats %>% 
    filter(bioregion %in% bioregions) 
  
  if(nrow(bioregion_test) < 2){
    if(nrow(bioregion_test) > 0){
      missing_bioreg <- subset(bioregions, !grepl(pattern = bioregion_test$bioregion, bioregions))
      spec_stats <- spec_stats %>% 
        ungroup() %>% 
        add_row(gbfp_SN = input$species, bioregion = missing_bioreg, utm1 = NA, utm2 = NA, utm5 = NA, utm10 = NA, sac_perc = NA)
    }else{
      for(bio in bioregions){
        spec_stats <- spec_stats %>% 
          ungroup() %>% 
          add_row(gbfp_SN = input$species, bioregion = bio, utm1 = NA, utm2 = NA, utm5 = NA, utm10 = NA, sac_perc = NA)
      }
    }
  }
  
  temp_stats_3 <- spec_stats  %>% 
    ungroup() %>% 
    filter(!is.na(bioregion)) %>% 
    select(-gbfp_SN) %>% 
    rename(`UTM 1x1km` = utm1,
           `UTM 2x2km` = utm2,
           `UTM 5x5km` = utm5,
           `UTM 10x10km` = utm10,
           `% UTM 1x1km squares with SACs*` = sac_perc) %>% 
    t()
  
  colnames(temp_stats_3) <- temp_stats_3[1,]
  temp_stats_3 <- temp_stats_3[-1,]
  temp_stats_3 <- as.data.frame(temp_stats_3) %>% 
    rownames_to_column(var = "grid") 
  print(temp_stats_3)
})
```

Emergence Graphs 
==========================================

```{r TrIAS constants, echo=FALSE, context="setup"}
evaluation_years <- maxjaar
```

```{r include=FALSE, context="server"}
renderPrint({
  evaluation_years <- input$evaluation_years
  print(evaluation_years)
})
```

Row {data-height=50}
----------------------------------

### Occupancy

```{r, context="server"}
renderPlot({
  
  spec_ind_sub <- spec_ind %>% 
    filter(gbifapi_acceptedScientificName == input$species)
  
  df_key <- df_ts_compact %>% 
    filter(taxonKey %in% spec_ind_sub$gbifapi_acceptedKey & year <= maxjaar)
  
  gam_occupancy <- apply_gam(
    df = df_key,
    y_var = "ncells",
    eval_years = evaluation_years,
    type_indicator = "occupancy",
    verbose = TRUE,
    saveplot = FALSE,
    y_label = "occupancy (km2)"
  )
  
  print(gam_occupancy$plot)
  
})
```

### Occupancy Corrected

```{r, context="server"}
renderPlot({
  
  spec_ind_sub <- spec_ind %>% 
    filter(gbifapi_acceptedScientificName == input$species)
  
  df_key <- df_ts_compact %>% 
    filter(taxonKey %in% spec_ind_sub$gbifapi_acceptedKey & year <= maxjaar)
  
  gam_occupancy_corrected <- apply_gam(
    df = df_key,
    y_var = "ncells",
    eval_years = evaluation_years,
    type_indicator = "occupancy",
    baseline_var = "c_ncells",
    verbose = TRUE,
    saveplot = FALSE,
    y_label = "occupancy (km2)"
  )
  
  print(gam_occupancy_corrected$plot)
  
})
```

Row {data-height=50}
----------------------------------

### Observations

```{r, context="server"}
renderPlot({
  
  spec_ind_sub <- spec_ind %>% 
    filter(gbifapi_acceptedScientificName == input$species)
  
  df_key <- df_ts_compact %>% 
    filter(taxonKey %in% spec_ind_sub$gbifapi_acceptedKey & year <= maxjaar)
  
  gam_observations <- apply_gam(
    df = df_key,
    y_var = "obs",
    eval_years = evaluation_years,
    type_indicator = "observations",
    verbose = TRUE,
    saveplot = FALSE
  )
  
  print(gam_observations$plot)
  
})
```

### Observations corrected

```{r, context="server"}
renderPlot({
  
  spec_ind_sub <- spec_ind %>% 
    filter(gbifapi_acceptedScientificName == input$species)
  
  df_key <- df_ts_compact %>% 
    filter(taxonKey %in% spec_ind_sub$gbifapi_acceptedKey & year <= maxjaar)
  
  gam_observations_corrected <- apply_gam(
    df = df_key,
    y_var = "obs",
    eval_years = evaluation_years,
    type_indicator = "observations",
    baseline_var = "cobs",
    verbose = TRUE,
    saveplot = FALSE
  )
  
  print(gam_observations_corrected$plot)
  
})
```

Emergence Graphs - SAC's
==========================================

Emergence Graphs - SPA's
==========================================

About
==========================================

Column {.tabset}
-------------------------------------

### T0 - Baseline

### T1 - Report

### Emergence Graphs

-!!placeholder TrIAS link!!- 

### Emergence Graphs - SAC's

### Emergence Graphs - SPA's

#### Basic

#### Corrected

### Partners
```{r logos}
renderImage({
  filename <- here("Image",
                   "Logo",
                   "INSTvoorNatuur_eng.jpeg")
  
  # Return a list containing the filename and alt text
  inbo_logo <- list(src = filename,
                    alt = paste("INBO - Logo"))
})
```